---
output: html_document
---
# This is the data merging and checking file

This is file that loads all the raw data from the Jibbon root trait project.  Then it stickes the datasets together and does a bunch of checks looking for:

1. Missing data
2. Duplicate data
3. Outliers
4. Spelling mistakes


Loading all the data
```{r,echo=TRUE,message=FALSE}
library(tidyverse)
air_porosity <- read_csv("data/air_porosity_complete.csv")
cluster_root_weight <- read_csv("data/cluster_root_weight_data_complete.csv")
dry_root_weight_1 <- read_csv("data/tissue_density_SRL_1.csv")
dry_root_weight_2 <- read_csv("data/tissue_density_SRL_2.csv")
root_length_data <- read_csv("data/branching_intensity_srl_lenght_complete.csv")
```


assemble mass data from two different files and make a column in the mass data set that is the unique id for matching
```{r}
bind_rows(dry_root_weight_1,dry_root_weight_2) %>%
  mutate(unique_id=paste0(site,sample,species,replicate)) %>%
  gather(key=type_of_data,value = mass_measurement,SRL,TD) -> mass_data_gathered
```

Are there any duplicates for SRL?
```{r}
srl.mass.data<-filter(mass_data_gathered, type_of_data=="SRL", !is.na(mass_measurement) )
head(sort(table(srl.mass.data$unique_id),decreasing = TRUE))
out<-filter(srl.mass.data,unique_id=="S20S2Ptilothrix deusta2")
```

Are there any duplicates for TD?
```{r}
td.mass.data<-filter(mass_data_gathered, type_of_data=="TD", !is.na(mass_measurement) )
head(sort(table(td.mass.data$unique_id),decreasing = TRUE))
out<-filter(td.mass.data,unique_id=="S3S3Leptocarpus tenax3")
```

make a column in the root length set that is the unique id for matching and check for duplicates


```{r}
root_length_data$unique_id<-paste0(root_length_data$site,root_length_data$sample,root_length_data$species,root_length_data$replicate)
sum(table(root_length_data$unique_id)>1)
```

Unique IDs in mass data that don't match the root length data

```{r}
srl.mass.data$unique_id[!srl.mass.data$unique_id %in% root_length_data$unique_id]
```


Things in root length data that don't match the mass data.

```{r}
root_length_data$unique_id[!root_length_data$unique_id %in% srl.mass.data$unique_id]
```

Then we match everything up and calculate SRL

```{r,message=FALSE,eval=TRUE,echo=TRUE}
inner_join(srl.mass.data,root_length_data,by="unique_id") %>%
  mutate(specific_root_length=Length_cm/mass_measurement) -> SRL.DF
```


```{r,message=FALSE,eval=FALSE,echo=FALSE}
library(plotly)
gg<-ggplot(SRL.DF,aes(x=specific_root_length,fill=species_code.x))+geom_histogram()+scale_x_log10()
gg
ggplotly(gg)

```

Checking the outliers

```{r,echo=TRUE,eval=TRUE}
arrange(SRL.DF,desc(specific_root_length)) %>%
  select(unique_id=unique_id,mass=mass_measurement,Length_cm=Length_cm,srl=specific_root_length)->SRL.DF.ORDERED

SRL.DF.ORDERED[1:10,]

arrange(SRL.DF,specific_root_length) %>%
  select(unique_id=unique_id,mass=mass_measurement,Length_cm=Length_cm,srl=specific_root_length)->SRL.DF.ORDERED

SRL.DF.ORDERED[1:10,]
```

Let's calculate density. 

Checking for duplicate samples

```{r}
air_porosity$unique_id<-paste0(air_porosity$site,air_porosity$sample,air_porosity$species,air_porosity$replicate)
out<-sort(table(air_porosity$unique_id),decreasing = TRUE)
out[out>1]
out<-filter(air_porosity,unique_id=="S14S2Lepidosperma neesii2")
```

Checking the matching from

```{r}
air_porosity$unique_id[!air_porosity$unique_id %in% td.mass.data$unique_id]

td.mass.data$unique_id[!td.mass.data$unique_id %in% air_porosity$unique_id]

```


Calculate density:

```{r}
inner_join(air_porosity,td.mass.data,by=c("site", "sample","species")) %>%
  select(site,sample,species,root_volume,mass_measurement) %>%
  mutate(density=mass_measurement/root_volume)-> density.df
```


Check for missing values either for volume or for mass
```{r}
density_missing_data<-filter(density.df,is.na(density))
```


```{r}
gg<-ggplot(density.df,aes(x=density,fill=species))+geom_histogram()+scale_x_log10()
gg
```


Checking for outliers:

```{r}
arrange(density.df,density) 

arrange(density.df,desc(density)) 
```


### Air porosity

```{r}
filter(air_porosity,air_porosity>0) %>%
ggplot(aes(y=air_porosity,x=species))+geom_point()

filter(air_porosity,air_porosity>0) %>%
  group_by(species) %>%
  summarize(mean(air_porosity))%>%
  knitr::kable(digits=2)
```

Possible ones that are crazy

```{r}
filter(air_porosity,air_porosity>35) 
```


### Number of tips

```{r}
root_length_data<- mutate(root_length_data,rbi=Number_of_tips/Length_cm)

filter(root_length_data,rbi>7) 


ggplot(root_length_data,aes(y=rbi,x=species))+geom_point()
filter(root_length_data,rbi<7) %>%
  group_by(species) %>%
  summarize(mean.rbi=mean(rbi))%>%
  arrange(mean.rbi)%>%
  knitr::kable(digits=2)
```

## Calulating median soil moisture values by species

```{r}
soil_moisture<-read_csv("data/Soil_Moisture.csv")
test<-inner_join(soil_moisture,mass_data_gathered)

group_by(test,species)%>%
  summarize(median.OM=median(OM),cv.OM=sd(OM)/mean(OM))%>%
  arrange(median.OM) -> species_median_OM
```

## Assmbling the mega_csv

```{r}
srl.for.joining<-select(SRL.DF,site=site.x,sample=sample.x,species=species.x,srl=specific_root_length,unique_id)
density.df.for.joining<-select(density.df,site,sample,species,density)
root.length.data.for.joining<-select(root_length_data,site,sample,species,rbi)
air.porosity.for.joining<-select(air_porosity,site,sample,species,air_porosity)
mean.soil<-summarize(group_by(soil_moisture,site),som=mean(OM))

full_join(density.df.for.joining,root.length.data.for.joining) %>%
   full_join(srl.for.joining) %>%
  full_join(air.porosity.for.joining) %>%
    full_join(mean.soil) ->mass_df

  write_csv(mass_df,"mass_output.csv")
  
  filter(mass_df,air_porosity>0)%>%
  ggplot(aes(x=som,y=air_porosity))+geom_point()+geom_smooth(method = "lm",se=FALSE)+scale_x_log10()->gg
  ggplotly(gg)
  
  ggplot(aes(x=som,y=air_porosity))+geom_point()+geom_smooth(method = "lm",se=FALSE)+scale_x_log10()->gg
  ggplotly(gg)
  
```

