---
output: html_document
---
# This is the data merging and checking file

This is file that loads all the raw data from the Jibbon root trait project.  Then it stickes the datasets together and does a bunch of checks looking for:

1. Missing data
2. Duplicate data
3. Outliers
4. Spelling mistakes


Loading all the data
```{r,echo=TRUE,message=FALSE}
library(tidyverse)
air_porosity<-read_csv("data/air_porosity_complete.csv")
cluster_root_weight<-read_csv("data/cluster_root_weight_data_complete.csv")
dry_root_weight_1<-read_csv("data/tissue_density_SRL_1.csv")
dry_root_weight_2<-read_csv("data/tissue_density_SRL_2.csv")
root_length_data<-read_csv("data/branching_intensity_srl_lenght_complete.csv")
```


assemble mass data from two different files
```{r}
mass_data<-rbind(dry_root_weight_1,dry_root_weight_2)
```

make a column in the mass data set that is the unique id for matching
```{r}
mass_data$unique_id<-paste0(mass_data$site,mass_data$sample,mass_data$species,mass_data$replicate)
head(sort(table(mass_data$unique_id),decreasing = TRUE))
filter(mass_data,unique_id=="S18S1Banksia oblongifolia1")
```

make a column in the root length set that is the unique id for matching and check for duplicates

```{r}
root_length_data$unique_id<-paste0(root_length_data$site,root_length_data$sample,root_length_data$species,root_length_data$replicate)
head(sort(table(root_length_data$unique_id),decreasing = TRUE))
filter(root_length_data,unique_id=="S20S1Plinthanthesis paradoxa1")
```

Unique IDs in mass data that don't match the root length data

```{r}
mass_data$unique_id[!mass_data$unique_id %in% root_length_data$unique_id]
```


Things in root length data that don't match the mass data.

```{r}
root_length_data$unique_id[!root_length_data$unique_id %in% mass_data$unique_id]
```

Then we match everything up and calculate SRL

```{r}
full_join(mass_data,root_length_data,by="unique_id") %>%
  mutate(specific_root_length=Length_cm/SRL) -> SRL.DF

library(plotly)
gg<-ggplot(SRL.DF,aes(x=specific_root_length,fill=species_code.x))+geom_histogram()+scale_x_log10()
ggplotly(gg)

gg<-ggplot(SRL.DF,aes(y=specific_root_length,x=species_code.x))+geom_violin()+scale_y_log10()
gg
```

Let's calculate density

```{r}
air_porosity$unique_id<-paste0(air_porosity$site,air_porosity$sample,air_porosity$species,air_porosity$replicate)
out<-sort(table(air_porosity$unique_id),decreasing = TRUE)
out[out>1]
```
